---
title: "Umesh_R_Project_readmitted_patient"
author: "Umesh B sangule"
date: "December 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Problem statement

Develop a statistical model to predict Patient Readmission using the dataset given "readmitted" is the dependent variable that needs to be predicted.

```{r}
library(dplyr)
library(corrplot)
library(ggplot2)
library(caret)
library(kernlab)
library(rpart)
library(randomForest)
library(e1071)
```

#  Fetching the dataset

```{r}
diabetes <- read.csv("diabetes_allcols.csv",sep = ",", header = T, na.strings = "?")
str(diabetes)
summary(diabetes)
```

For getting the readmission prediction we dont need many columns hence we should be selecting only relevant values to make it simpler.Also we need to remove those columns which mainly has NA values.

```{r}
data <- select(diabetes,  -encounter_id, -patient_nbr, -weight,-(25:41),-(43:47))
head(data)
```

Basic summary of selected data and correlation plot of numeric data....

```{r}
summary(data)

numeric_data <- select_if(data, is.numeric)
c <- cor(numeric_data, use= "pairwise.complete.obs")

corrplot(c)
```

```{r}
data$race[is.na(data$race)] <- "Other"
```

Relation of different variables....

```{r}
plot(data$age, main = "age distribution") # age: mode 70-80yrs normal distribution, left skewed
plot(data$gender, main = "gender distribution") # gender: female 53% male 47%
plot(data$A1Cresult, main = "A1C") # A1Cresult: 84% no A1c results, 8% >8
plot(data$readmitted, main = "readmissions") # readmission: >50% no readmission
plot(data$admission_source, main = "admission source") # emergency 60%
```

# Changing categorical variables

```{r}
data2 <- data
data2$diag_1 <- as.numeric(levels(data2$diag_1)[data2$diag_1])
data2$diag_2 <- as.numeric(levels(data2$diag_2)[data2$diag_2])
data2$diag_3 <- as.numeric(levels(data2$diag_3)[data2$diag_3])

```

# Diagnosis1(diag_1) variable categorization....

```{r}
data2$diagnosis_group <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels =     c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                   "Musculoskeletal","Genitourinary","Neoplasms","other"))

data2$diagnosis_group[data2$diag_1 >= 390 &
                        data2$diag_1 <= 459 | data2$diag_1 == 785] <- "circulatory"

data2$diagnosis_group[data2$diag_1 >= 460 & 
                        data2$diag_1 <= 519 | data2$diag_1 == 786] <- "respiratory"

data2$diagnosis_group[data2$diag_1 >= 520 & 
                        data2$diag_1 <= 579 | data2$diag_1==787] <- "Digestive"

data2$diagnosis_group[data2$diag_1>=250 & data2$diag_1 < 251] <- "Diabetes"

data2$diagnosis_group[data2$diag_1>800 & data2$diag_1 <= 999] <- "Injury"

data2$diagnosis_group[data2$diag_1>=710 & data2$diag_1 <= 739] <- "Musculoskeletal"

data2$diagnosis_group[data2$diag_1>=580 & 
                        data2$diag_1 <= 629 | data2$diag_1==788] <- "Genitourinary"

data2$diagnosis_group[data2$diag_1>=140 & data2$diag_1 <= 239 | data2$diag_1>=790 & 
                        data2$diag_1 <= 799 | data2$diag_1==780 | data2$diag_1>=240 & data2$diag_1 < 250 |
                        data2$diag_1>=251 & data2$diag_1 <= 279 | data2$diag_1>=680 & data2$diag_1 <= 709 |
                        data2$diag_1>=001 & data2$diag_1 <= 139 | data2$diag_1==781 |
                      data2$diag_1==782 | data2$diag_1==784] <- "Neoplasms"
```

# Diagnosis2(diag_2) variable categorization....

```{r}
data2$diagnosis_2 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_2[data2$diag_2>=390 & data2$diag_2 <= 459 | data2$diag_2==785] <- "circulatory"
data2$diagnosis_2[data2$diag_2>=460 & data2$diag_2 <= 519 | data2$diag_2==786] <- "respiratory"
data2$diagnosis_2[data2$diag_2>=520 & data2$diag_2 <= 579 | data2$diag_2==787] <- "Digestive"
data2$diagnosis_2[data2$diag_2>=250 & data2$diag_2 < 251] <- "Diabetes"
data2$diagnosis_2[data2$diag_2>800 & data2$diag_2 <= 999] <- "Injury"
data2$diagnosis_2[data2$diag_2>=710 & data2$diag_2 <= 739] <- "Musculoskeletal"
data2$diagnosis_2[data2$diag_2>=580 & data2$diag_2 <= 629 | data2$diag_2==788] <- "Genitourinary"
data2$diagnosis_2[data2$diag_2>=140 & data2$diag_2 <= 239 | data2$diag_2>=790 & 
                        data2$diag_2 <= 799 | data2$diag_2==780 | data2$diag_2>=240 & data2$diag_2 < 250 |
                        data2$diag_2>=251 & data2$diag_2 <= 279 | data2$diag_2>=680 & data2$diag_2 <= 709 |
                        data2$diag_2>=001 & data2$diag_2 <= 139 | data2$diag_2==781 |
                        data2$diag_2==782 | data2$diag_2==784] <- "Neoplasms"
```

# Diagnosis3(diag_3) variable categorization....

```{r}
data2$diagnosis_3 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_3[data2$diag_3>=390 & data2$diag_3 <= 459 | data2$diag_3==785] <- "circulatory"
data2$diagnosis_3[data2$diag_3>=460 & data2$diag_3 <= 519 | data2$diag_3==786] <- "respiratory"
data2$diagnosis_3[data2$diag_3>=520 & data2$diag_3 <= 579 | data2$diag_3==787] <- "Digestive"
data2$diagnosis_3[data2$diag_3>=250 & data2$diag_3 < 251] <- "Diabetes"
data2$diagnosis_3[data2$diag_3>800 & data2$diag_3 <= 999] <- "Injury"
data2$diagnosis_3[data2$diag_3>=710 & data2$diag_3 <= 739] <- "Musculoskeletal"
data2$diagnosis_3[data2$diag_3>=580 & data2$diag_3 <= 629 | data2$diag_3==788] <- "Genitourinary"
data2$diagnosis_3[data2$diag_3>=140 & data2$diag_3 <= 239 | data2$diag_3>=790 & 
                        data2$diag_3 <= 799 | data2$diag_3==780 | data2$diag_3>=240 & data2$diag_3 < 250 |
                        data2$diag_3>=251 & data2$diag_3 <= 279 | data2$diag_3>=680 & data2$diag_3 <= 709 |
                        data2$diag_3>=001 & data2$diag_3 <= 139 | data2$diag_3==781 |
                        data2$diag_3==782 | data2$diag_3==784] <- "Neoplasms"

```

```{r}
summary(data2)
```

# Admission_source variable categorization

```{r}

data2$admission_source <- factor( rep("other", nrow(data2)), ordered = F, 
                             levels = c("clinic_referral", "emergency","other"))

data2$admission_source[data2$admission_source_id == c(1,2,3)]<- "clinic_referral"

data2$admission_source[data2$admission_source_id == 7]<- "emergency"

head(data2$admission_source)
```

# Discharged_to variable categorization.....

```{r}

data2$discharged_to <- factor( rep("transferred", nrow(data2)), ordered = F, 
                                  levels = c("home", "transferred", "left_AMA"))

data2$discharged_to[data2$discharge_disposition_id==c(1,6,8)]<- "home"

data2$discharged_to[data2$discharge_disposition_id==7]<- "left_AMA"

data2 <- select(data2, -diag_1, -diag_2, -diag_3, -admission_type_id, -discharge_disposition_id)

data2 <- select(data2, -medical_specialty)

data2 <- rename(data2, diag1 = diagnosis_group, diag2=diagnosis_2, diag3 = diagnosis_3)
summary(data2)
```

# payer_code variable categorization.......

```{r}
data2$payer_code2 <- factor( rep("other", nrow(data2)),
                             ordered = F, levels = c("other", "self_pay"))

data2$payer_code2[data2$payer_code=="SP"]<- "self_pay"

data2 <- select(data2, -payer_code)
data2 <- select(data2, -admission_source_id)
data2 <- rename(data2, payer_code=payer_code2)
summary(data2)
head(data2)
```

PCA Principal Component Analysis
In order to reduce 23 dimensions or knowing the significant impact of our reduction we will do PCA.

## PCA (principal Component Analysis)

```{r}
numeric_data <- scale(numeric_data)
pcaObj <- princomp(numeric_data, cor = TRUE, scores = TRUE, covmat = NULL)
summary(pcaObj)
print(pcaObj)
names(pcaObj)
plot(pcaObj)
pcaObj$loadings
biplot(pcaObj, cex = 0.7)
final_data <- as.data.frame(pcaObj$scores)
```

# Converting the dataset into training data and validation data.

```{r}
set.seed(123)
inTrain <- createDataPartition(y = data2$readmitted, p = .67,list = FALSE)
train <- data2[ inTrain,]
test <- data2[-inTrain,]
nrow(train)
nrow(test) 
summary(train)
summary(test)
```

## Regression 

Since the output variable readmitted is a categorical variable we need to find logistic regression 
```{r}
train_nonbinary <- train
test_nonbinary <- test

train$readmitted <- ifelse(train$readmitted == train$readmitted[1], 0 , 1)
test$readmitted <- ifelse(test$readmitted == test$readmitted[1], 0 , 1)
head(test)

logit_model <- glm(readmitted ~., data = train, family = binomial(link = 'logit'))

summary(logit_model)

plot(logit_model)

pred_logit <- predict(logit_model,test, type = "response")
#pred_logit

pred_logit <- ifelse(pred_logit > 0.5, 1, 0)
#pred_logit

result <- as.data.frame(table(pred_logit,test$readmitted))
#result

CorrectlyPredicted <- result[1,3]+result[4,3]

accuracy <- CorrectlyPredicted/nrow(test)
accuracy

senstivity_result <- result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result <- result[1,3]/(result[3,3]+result[1,3])
specificity_result
```

Adding dimension which are most significant and then doing logistic regression

```{r}

normal_fit <- glm(readmitted ~ race + age + discharged_to + time_in_hospital +
             num_lab_procedures + num_procedures + num_medications + number_outpatient +
             number_emergency + number_inpatient + number_diagnoses +
             insulin + change + diabetesMed + diag1 + diag2 + diag3 + A1Cresult, 
             data = train, family = binomial(link = 'logit'))

summary(normal_fit)

pred_logit <- predict(normal_fit,test, type = "response")
#pred_logit

pred_logit <- ifelse(pred_logit > 0.5, 1, 0)
#pred_logit

result <- as.data.frame(table ( pred_logit , test$readmitted ))
result

CorrectlyPredicted <- result[1, 3] + result[4, 3]

accuracy <- CorrectlyPredicted / nrow(test)
accuracy

senstivity_result <- result[4, 3] / (result[2, 3] + result[4, 3])
senstivity_result

specificity_result <- result[1, 3] / (result[3, 3] + result[1, 3])
specificity_result

```


## Rpart Decision Tree ##

R part decision Trees Prediction

Prediction using three level category

Training

```{r}
# rpart_tree <- rpart(formula = readmitted ~ age + discharged_to + time_in_hospital +
 #                     num_lab_procedures+num_procedures+num_medications  +         
 # number_outpatient + number_emergency + number_inpatient + number_diagnoses +
 #                     insulin + change + diabetesMed + diag1 + diag2 + diag3 + A1Cresult, 
 #                   data=train_nonbinary, method = 'class')
# summary(rpart_tree)
```

Prediction 

```{r}
# pred_tree <- predict(rpart_tree, test_nonbinary, type="class")
# head(pred_tree)
```

Performance

```{r}
#table(predict(rpart_tree, test_nonbinary, type="class"), test_nonbinary$readmitted)

# result <- as.data.frame(table(predict(rpart_tree, test_nonbinary, type="class"), 
#                            test_nonbinary$readmitted))
#  result

# confusionMatrix(pred_tree, test_nonbinary$readmitted)

# prop.table(table(test_nonbinary$readmitted, pred_tree),1)
```

Prediction in binary format

Training

```{r}
# train$readmitted <- ifelse(train$readmitted == train$readmitted[1], 0, 1)
# test$readmitted <- ifelse(test$readmitted == test$readmitted[1], 0, 1)

# rpart_tree <- rpart(formula = readmitted ~ age + discharged_to + time_in_hospital +
#                      num_lab_procedures+num_procedures+num_medications  +         number_outpatient + number_emergency + number_inpatient + number_diagnoses +
#                      insulin + change + diabetesMed + diag1 + diag2 + diag3 + A1Cresult, 
#                    data=train, method = 'class')

 # summary(rpart_tree)
```

Prediction

```{r}
#pred_tree <- predict(rpart_tree, test, type="class")

```

Performance

```{r}
#table(predict(rpart_tree, test, type = "class"), test$readmitted)
#result <- as.data.frame(table(predict(rpart_tree, test, type = "class"), test$readmitted))
#result

#CorrectlyPredicted <- result[1, 3] + result[4, 3]

#accuracy <- CorrectlyPredicted / nrow(test)
#accuracy

#senstivity_result <- result[4, 3] / (result[2, 3] + result[4, 3])
#senstivity_result

#specificity_result <- result[1, 3] / (result[3, 3] + result[1, 3])
#specificity_result

#prop.table(table(test$readmitted, pred_tree), 1)
```

